¥documentclass[11pt]{amsart}

¥usepackage[T1]{fontenc}
¥usepackage[utf8]{inputenc}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage{hyperref}

¥newcommand{¥LL}{¥Lambda}
¥newcommand{¥OO}{¥mathbb{O}}
¥newcommand{¥ZZ}{¥mathbb{Z}}
¥newcommand{¥RR}{¥mathbb{R}}

¥newtheorem*{observation}{Observation}
¥newtheorem{example}{Example}[section]

¥title[A companion note on the $¥sigma$-twist]%
{A companion note¥¥
¥large Kirmse--Coxeter and the $¥sigma$-twist, by concrete example}
¥author{Jens K¥"oplinger}
¥date{April 2026}

¥begin{document}
¥maketitle

¥begin{abstract}
This note accompanies the paper ``An order on the Leech lattice from
transposition-twisted octonions.''  Its purpose is purely expository:
to walk through the construction step by step, pairing every abstract
definition with a concrete hand-checkable example, and to do so first
at the‾$E_8$ level (where Coxeter's correction of Kirmse provides the
historical precedent) and then at the Leech level (where the
$¥sigma$-twist plays the analogous role).  No new results are
proved.  Readers who are comfortable with the main paper will find
nothing new here; readers who want to see every piece verified by
hand will find the explicit numerics they may have missed.
¥end{abstract}

% ===================================================================
¥section{Preliminaries}
% ===================================================================

This note follows the conventions of the main paper.  We work with
the real octonion algebra‾$¥OO$ on $¥RR^8$ with basis
$¥{e_0, e_1, ¥ldots, e_7¥}$, and with Wilson's construction of the
Leech lattice inside the triple octonion space
$¥RR^{24} = ¥OO_1 ¥oplus ¥OO_2 ¥oplus ¥OO_3$.

The guiding principle throughout is: ¥emph{distinguish abstract
structure from the specific lattice to which it is applied}.  The
same ambient algebra can host many lattices; the same lattice can
be tested against many bilinear products on the ambient space.
``Closure of this lattice under this product'' is a question about
both, not just one.

% ===================================================================
¥section{The $E_8$ lattice}¥label{sec:e8}
% ===================================================================

¥subsection*{Abstract}

The lattice‾$E_8$ is the unique even unimodular lattice in‾$¥RR^8$
(up to isomorphism).  ``Even'' means every vector‾$v ¥in E_8$ has
integer squared norm $¥|v¥|^2 ¥in 2¥ZZ$.  ``Unimodular'' means
$E_8$ equals its own dual lattice (the set of vectors having
integer inner product with every element of‾$E_8$).  It contains
$240$ vectors of minimal squared norm‾$2$, the ¥emph{roots}.

¥subsection*{Concrete}

Following Wilson, take for $L$ the integer span of the following
240 octonions:
¥begin{itemize}
¥item ¥textbf{112 ``type-1'' roots} of the form
  $¥pm e_a ¥pm e_b$ with $0 ¥le a < b ¥le 7$;
¥item ¥textbf{128 ``type-2'' roots} of the form
  $¥tfrac{1}{2}(¥varepsilon_0 e_0 + ¥varepsilon_1 e_1 + ¥cdots
  + ¥varepsilon_7 e_7)$ where each
  $¥varepsilon_i ¥in ¥{+1, -1¥}$ and the number of‾$-1$s is odd.
¥end{itemize}

Equivalently,
¥[
  L = ¥{v ¥in ¥ZZ^8 : ¥sum_i v_i ¥text{ is even}¥}
   ¥;¥cup¥;
      ¥{v ¥in (¥ZZ + ¥tfrac{1}{2})^8 : ¥sum_i v_i ¥text{ is odd}¥}.
¥]

¥begin{example}
The vector $(1, 1, 0, 0, 0, 0, 0, 0) = e_0 + e_1$ is a type-1 root:
integer coordinates, sum‾$2$.  The vector
$¥tfrac{1}{2}(-1, 1, 1, 1, 1, 1, 1, 1)$ is a type-2 root:
half-integer coordinates, one $-1$ (odd count), coordinate sum $3$
(odd).  Both lie in‾$L$; both have squared norm‾$2$.
¥end{example}

¥begin{example}
Distinguished element in‾$L$ to be used throughout:
¥[
  s ¥;=¥; ¥tfrac{1}{2}(-e_0 + e_1 + e_2 + e_3 + e_4 + e_5 + e_6 + e_7),
¥]
of squared norm‾$2$.  In coordinates $s = (-¥tfrac{1}{2},
¥tfrac{1}{2}, ¥tfrac{1}{2}, ¥tfrac{1}{2}, ¥tfrac{1}{2}, ¥tfrac{1}{2},
¥tfrac{1}{2}, ¥tfrac{1}{2})$: half-integer, one‾$-1/2$ (odd count),
sum‾$3$ (odd).  Confirmed in‾$L$.
¥end{example}

% ===================================================================
¥section{The octonion algebra}¥label{sec:oct}
% ===================================================================

¥subsection*{Abstract}

The octonions‾$¥OO$ are the unique real ¥emph{normed division
algebra} of dimension‾$8$: a real $8$-dimensional vector space with
a bilinear product‾$x ¥cdot y$ and a positive-definite quadratic
norm satisfying $¥|x ¥cdot y¥| = ¥|x¥| ¥cdot ¥|y¥|$.  The algebra is
non-commutative and non-associative (but alternative).

¥subsection*{Concrete}

Fix the basis $¥{e_0, e_1, ¥ldots, e_7¥}$ with $e_0$ the
multiplicative identity.  The multiplication on the imaginary units
is determined by the seven ¥emph{Fano triples}
¥begin{equation}¥label{eq:fano}
  (1,2,4),¥; (2,3,5),¥; (3,4,6),¥; (4,5,7),¥; (5,6,1),¥; (6,7,2),¥;
  (7,1,3),
¥end{equation}
read cyclically: for $(a, b, c)$ above, $e_a ¥cdot e_b = +e_c$,
$e_b ¥cdot e_c = +e_a$, $e_c ¥cdot e_a = +e_b$, with opposite signs
for the reversed products.  Every imaginary unit squares to‾$-e_0$.

¥begin{example}
From the triple‾$(1,2,4)$: $¥,e_1 ¥cdot e_2 = +e_4$ and
$e_2 ¥cdot e_1 = -e_4$.  From‾$(7,1,3)$: $¥,e_1 ¥cdot e_3 = +e_7$.
From‾$(6,7,2)$: $¥,e_7 ¥cdot e_2 = +e_6$, so
$e_2 ¥cdot e_7 = -e_6$.
¥end{example}

¥subsection*{Closure of $L$ under the octonion product}

Under the Fano convention ¥eqref{eq:fano}, the lattice‾$L$ is closed
under octonion multiplication: for every $u, v ¥in L$, the product
$u ¥cdot v ¥in L$.  This is a classical result of Coxeter‾(1946),
verified in the main paper by computing all $8 ¥times 8 = 64$
products of a $¥ZZ$-basis of‾$L$ in exact arithmetic and checking
that each lies again in‾$L$.

¥begin{example}
A sample product.  From $s ¥cdot s$ with $s$ as above, using the
identity $(¥sum_{k=1}^{7} e_k)^2 = -7 e_0$:
¥[
  s ¥cdot s ¥;=¥;
    -¥tfrac{3}{2}¥, e_0 - ¥tfrac{1}{2}(e_1 + e_2 + e_3 + e_4 + e_5
    + e_6 + e_7).
¥]
Its coordinates are all half-integer; the coordinate sum is
$-¥tfrac{3}{2} - 7 ¥cdot ¥tfrac{1}{2} = -5$, odd.  Hence
$s ¥cdot s ¥in L$.
¥end{example}

% ===================================================================
¥section{The Kirmse--Coxeter precedent}¥label{sec:kirmse}
% ===================================================================

¥subsection*{Historical remark}

Kirmse (1924) proposed a specific set of 240 integral octonions
together with a multiplication table.  He asserted the set was
closed under the multiplication.  Coxeter (1946) noticed a subtle
error: Kirmse's multiplication table, as written, did not produce
closure.  Coxeter showed that if one permutes two basis indices in
the Fano-plane labelling, one obtains a (different) multiplication
table under which the 240-element set ¥emph{is} closed.  There are
seven such Coxeter-corrected conventions, one for each pair of
imaginary indices; each gives one of the seven maximal orders of
the integral octonions discussed in Conway--Smith.

The Fano triples ¥eqref{eq:fano} used in this note and in the main
paper are one of Coxeter's seven corrected conventions.  Under this
convention $L$ closes, as stated above.  We do not reproduce
Kirmse's original table here; the point of recalling the story is
the structural pattern, not the historical arithmetic:

¥begin{observation}
A convention of octonion multiplication fixes how basis products
are assigned.  A different convention, obtained by transposing two
basis indices, assigns products differently.  For some lattices, the
two conventions are interchangeable (the lattice is closed under
both); for others, only one convention closes.  Which lattices
behave which way is precisely the question that Kirmse got wrong
and Coxeter got right.
¥end{observation}

¥subsection*{A concrete lattice where the story does matter}

For the lattice‾$L$ itself, this distinction is invisible: $L$ is
symmetric under every coordinate permutation, so transposing two
basis indices produces an algebra under which $L$ also closes.  To
see the Kirmse--Coxeter phenomenon concretely at the $E_8$ level we
need a lattice that is ¥emph{not} symmetric under such a swap.  One
appears naturally in Wilson's Leech construction: the sublattice
¥[
  Ls ¥;=¥; ¥{¥ell ¥cdot s : ¥ell ¥in L¥},
¥]
the image of $L$ under right-multiplication by the element‾$s$
above.  This is again a full-rank sublattice of‾$L$, of index‾$16$
(see the main paper, Section‾2.3).  It is a lattice, so of course it
is closed under addition; the question is whether it is closed under
octonion multiplication.

¥begin{example}[Standard product fails on‾$Ls$]¥label{ex:Ls-fails}
Take
¥[
  a = (-1,¥, 0,¥, 0,¥, 0,¥, 1,¥, 0,¥, 1,¥, 1), ¥qquad
  b = (-1,¥, 1,¥, 0,¥, 0,¥, 0,¥, 1,¥, 0,¥, 1).
¥]
Both $a$ and $b$ are in $Ls$ (they appear as basis vectors of‾$Ls$
produced in the main paper's exact-arithmetic verification).  Their
octonion product, computed under the Fano convention
¥eqref{eq:fano}, is
¥[
  a ¥cdot b = (0,¥, -2,¥, 2,¥, 1,¥, -2,¥, -1,¥, -1,¥, -1).
¥]
This vector has integer coordinates and lies in‾$L$ --- but it is
¥emph{not} an integer combination of a $¥ZZ$-basis of‾$Ls$.  The
verification is carried out with exact rational arithmetic in the
script
¥href{../python_project/src/symbolic_proof_checks.py}{¥texttt{symbolic¥_proof¥_checks.py}}.

Consequently, $Ls ¥cdot Ls ¥not¥subseteq Ls$.
¥end{example}

This is the structural echo of Kirmse's situation at the sublattice
level: a specific sublattice, under a specific octonion product,
fails to close.

% ===================================================================
¥section{The Coxeter transposition, abstractly}¥label{sec:swap}
% ===================================================================

¥subsection*{Abstract}

Fix a transposition $¥sigma = (s¥; t)$ on the imaginary basis
indices $¥{1, ¥ldots, 7¥}$ (for example, $¥sigma = (1¥;2)$).
Extend‾$¥sigma$ linearly to a map $¥RR^8 ¥to ¥RR^8$ that fixes
$e_0$ and swaps $e_s ¥leftrightarrow e_t$.  This‾$¥sigma$ is an
¥emph{isometry} of‾$¥RR^8$: it preserves the Euclidean inner
product.  It is also an involution: $¥sigma^2 = ¥mathrm{id}$.

Starting from the standard product‾$¥cdot$, the map‾$¥sigma$ induces
a ¥emph{$¥sigma$-twisted product} on the same $¥RR^8$:
¥begin{equation}¥label{eq:twist}
  x ¥cdot_¥sigma y ¥;=¥; ¥sigma¥bigl(¥sigma(x) ¥cdot ¥sigma(y)¥bigr).
¥end{equation}
The pair $(¥RR^8, ¥cdot_¥sigma)$ is again an octonion algebra,
abstractly isomorphic to $(¥RR^8, ¥cdot)$ via‾$¥sigma$ itself.  But
it is a ¥emph{different bilinear product} on the same $¥RR^8$: it
assigns different specific products to pairs of basis vectors.

¥subsection*{Concrete}

Take $¥sigma = (1¥; 2)$, which swaps $e_1$ and‾$e_2$ (and fixes
$e_0, e_3, e_4, e_5, e_6, e_7$).  Then
¥begin{align*}
  e_1 ¥cdot_¥sigma e_2
    &= ¥sigma(¥sigma(e_1) ¥cdot ¥sigma(e_2))
     = ¥sigma(e_2 ¥cdot e_1)
     = ¥sigma(-e_4)
     = -e_4,
¥end{align*}
whereas $e_1 ¥cdot e_2 = +e_4$ under the standard product.  Similarly
one checks $e_1 ¥cdot_¥sigma e_4 = +e_2$ (vs.¥ $-e_2$ under
$¥cdot$), and so on: the two products disagree on some pairs of
basis elements.

¥subsection*{What is twisted and what is invariant}

¥begin{itemize}
¥item ¥textbf{Twisted:} the multiplication table.  Under
  $¥cdot_¥sigma$, the Fano triples ¥eqref{eq:fano} are replaced by
  $(¥sigma(a), ¥sigma(b), ¥sigma(c))$ for each original triple.
¥item ¥textbf{Invariant under‾$¥sigma$ (as a linear map):}
  the underlying vector space‾$¥RR^8$, the Euclidean inner product,
  the decomposition into integer and half-integer vectors, and the
  lattice‾$L$ itself ($¥sigma(L) = L$ since $¥sigma$ is a coordinate
  permutation).
¥item ¥textbf{Not invariant under‾$¥sigma$:} specific sublattices
  of‾$L$ that depend on a distinguished element, such as‾$Ls$.
  Since $Ls$ is defined as $¥{¥ell ¥cdot s : ¥ell ¥in L¥}$, applying
  $¥sigma$ gives
  ¥[
    ¥sigma(Ls) = ¥{¥sigma(¥ell) ¥cdot_{¥sigma} ¥sigma(s)
                    : ¥ell ¥in L¥}
    ¥;¥ne¥; Ls
  ¥]
  in general.  The main paper verifies $¥sigma(Ls) ¥ne Ls$
  directly: for $¥sigma = (1¥;2)$, the vector
  $(-1, 0, 1, 0, 0, 1, 0, 1)$ lies in‾$¥sigma(Ls)$ but not
  in‾$Ls$.
¥end{itemize}

The last two bullets together say: ¥emph{$¥sigma$ is an isometry of
$¥RR^8$ that fixes‾$L$ but moves‾$Ls$ to a distinct sublattice.}

% ===================================================================
¥section{The twist at the $E_8$ level changes the story}
% ===================================================================

The sublattice $¥sigma(Ls)$ is the image of‾$Ls$ under the
transposition.  It has the same rank as‾$Ls$, the same index in‾$L$,
and is isometric to‾$Ls$.  But its elements are different vectors.
The question of closure under the standard octonion product is now
posed afresh.

¥begin{example}[Standard product succeeds on‾$¥sigma(Ls)$]
¥label{ex:sigmaLs-succeeds}
A $¥ZZ$-basis of‾$¥sigma(Ls)$ is obtained by applying‾$¥sigma$
(coordinate swap of positions $1, 2$) to each basis vector of‾$Ls$.
Using the basis generated in the main paper's verification, all
$8 ¥times 8 = 64$ products of basis vectors of‾$¥sigma(Ls)$, under
the standard octonion product‾$¥cdot$, lie again in‾$¥sigma(Ls)$.
The verification is in
¥href{../python_project/src/symbolic_proof_checks.py}{¥texttt{symbolic¥_proof¥_checks.py}}
as Lemma‾D.

Hence $¥sigma(Ls) ¥cdot ¥sigma(Ls) ¥subseteq ¥sigma(Ls)$.
¥end{example}

¥subsection*{Two equivalent readings}

Because $¥sigma$ is an isometry and an algebra isomorphism
$(¥RR^8, ¥cdot) ¥to (¥RR^8, ¥cdot_¥sigma)$, one can restate
Example‾¥ref{ex:sigmaLs-succeeds} in an equivalent form on the
original lattice‾$Ls$ with the twisted product:
¥[
  Ls ¥cdot_¥sigma Ls ¥;¥subseteq¥; Ls.
¥]
(Apply $¥sigma$ to both sides of the inclusion
$¥sigma(Ls) ¥cdot ¥sigma(Ls) ¥subseteq ¥sigma(Ls)$ and use
$¥sigma^2 = ¥mathrm{id}$ together with
¥eqref{eq:twist}.)  The two forms are tautologically equivalent.

Reading both against Example‾¥ref{ex:Ls-fails}:
¥[
  ¥underbrace{Ls ¥cdot Ls ¥not¥subseteq Ls}_{¥text{standard product
  fails on $Ls$}}
  ¥qquad¥text{versus}¥qquad
  ¥underbrace{Ls ¥cdot_¥sigma Ls ¥subseteq Ls}_{¥text{twisted product
  succeeds on $Ls$}}.
¥]
Equivalently:
¥[
  ¥underbrace{Ls ¥cdot Ls ¥not¥subseteq Ls}_{¥text{$Ls$ fails}}
  ¥qquad¥text{versus}¥qquad
  ¥underbrace{¥sigma(Ls) ¥cdot ¥sigma(Ls) ¥subseteq ¥sigma(Ls)}
    _{¥text{$¥sigma(Ls)$ succeeds}}.
¥]

¥subsection*{Interpretation}

Two different bilinear products on‾$¥RR^8$ (namely‾$¥cdot$
and‾$¥cdot_¥sigma$) are isomorphic as abstract algebras.  Every
theorem phrased in the language of the abstract algebra carries
over.  But ``does lattice‾$X$ close under product‾$¥star$?'' is not
a theorem phrased purely in the abstract algebra: it also names a
specific subset $X ¥subset ¥RR^8$, and the answer depends on how the
isomorphism relates $X$ to itself.

If $¥sigma(X) = X$, the answer is the same for both products.  If
$¥sigma(X) ¥ne X$, the answer can flip: $X$ may fail to close
under‾$¥cdot$ while $¥sigma(X)$ closes under‾$¥cdot$; equivalently,
$X$ closes under $¥cdot_¥sigma$ while‾$¥sigma(X)$ fails under
$¥cdot_¥sigma$.  The lattice‾$L$ itself is $¥sigma$-invariant and
therefore closes under both products.  The sublattice‾$Ls$ is
¥emph{not} $¥sigma$-invariant, and this is exactly what opens the
gap between the two closure statements.

This is the structural lesson of Kirmse--Coxeter in isolation.  The
same lesson, applied one dimension higher, is the content of the
main paper.

% ===================================================================
¥section{Wilson's Leech lattice}¥label{sec:wilson-leech}
% ===================================================================

¥subsection*{Abstract}

The Leech lattice‾$¥LL$ is the unique even unimodular lattice of
rank‾$24$ with no roots (i.e.¥ no vectors of squared norm‾$2$).  Its
minimal vectors have squared norm‾$4$, and there are exactly
$196{,}560$ of them.

¥subsection*{Concrete (Wilson's construction)}

Identify $¥RR^{24}$ with $¥OO_1 ¥oplus ¥OO_2 ¥oplus ¥OO_3$: three
copies of the octonions, labelled $¥alpha = 1, 2, 3$.  Let
$¥bar{s} = ¥tfrac{1}{2}(-e_0 - e_1 - e_2 - ¥cdots - e_7)$ (the
octonion conjugate of the negative of the earlier‾$s$; it also lies
in‾$L$).  Define two sublattices of‾$L$:
¥[
  Ls = ¥{¥ell ¥cdot s : ¥ell ¥in L¥}, ¥qquad
  L¥bar{s} = ¥{¥ell ¥cdot ¥bar{s} : ¥ell ¥in L¥},
¥]
each of index‾$16$ in‾$L$.  Wilson's Leech lattice is then
¥begin{multline}¥label{eq:leech}
  ¥LL ¥;=¥; ¥bigl¥{(x, y, z) ¥in L^3 ¥;:¥¥
    x + y,¥; x + z,¥; y + z ¥;¥in¥; L¥bar{s}; ¥quad
    x + y + z ¥;¥in¥; Ls ¥bigr¥}.
¥end{multline}

The sublattice‾$Ls$ from Section‾¥ref{sec:kirmse} is precisely the
``condition-(3) block sum'' lattice in Wilson's definition.

% ===================================================================
¥section{The standard triple product on $¥RR^{24}$}
% ===================================================================

¥subsection*{Abstract}

The natural bilinear product on $¥RR^{24} = ¥OO_1 ¥oplus ¥OO_2
¥oplus ¥OO_3$ built from a single octonion product is a
¥emph{triple product with $¥ZZ_3$-symmetric cross-block routing}.
For $u = (x, y, z)$ and $v = (x', y', z')$ define
¥[
  u ¥star v ¥;=¥; (P_1, P_2, P_3),
¥]
where
¥begin{align*}
  P_1 &= x ¥cdot x' + z ¥cdot y' + y ¥cdot z', ¥¥
  P_2 &= y ¥cdot y' + x ¥cdot z' + z ¥cdot x', ¥¥
  P_3 &= z ¥cdot z' + y ¥cdot x' + x ¥cdot y'.
¥end{align*}
Same-block products stay in the block; cross-block products
$O_¥alpha ¥times O_¥beta$ land in $O_¥gamma$ with
$¥{¥alpha, ¥beta, ¥gamma¥} = ¥{1, 2, 3¥}$.  The three blocks play
the same role under cyclic rotation $1 ¥to 2 ¥to 3 ¥to 1$.

¥subsection*{Why condition‾(3) is the obstruction}

A direct computation from the definitions gives
¥[
  P_1 + P_2 + P_3 ¥;=¥; (x + y + z) ¥cdot (x' + y' + z').
¥]
Thus ``$u ¥star v$ satisfies Wilson's condition‾(3)'' amounts to:
``the product of two elements of‾$Ls$ lies in‾$Ls$''.

Using $u, v ¥in ¥LL$ means $x + y + z ¥in Ls$ and $x' + y' + z'
¥in Ls$, so the question reduces exactly to whether $Ls ¥cdot Ls
¥subseteq Ls$.  By Example‾¥ref{ex:Ls-fails}, it does not.  Hence:

¥begin{example}[Standard triple product fails on‾$¥LL$]
¥label{ex:triple-fails}
There exist $u, v ¥in ¥LL$ for which $u ¥star v ¥notin ¥LL$: the
block sum of $u ¥star v$ need not lie in‾$Ls$, failing Wilson's
condition‾(3).  The concrete obstruction is precisely the pair
$(a, b)$ of Example‾¥ref{ex:Ls-fails}: any $u, v ¥in ¥LL$ whose
triple sums equal‾$a, b$ respectively produce a $u ¥star v$ whose
block sum equals‾$a ¥cdot b ¥notin Ls$.
¥end{example}

% ===================================================================
¥section{The $¥sigma$-twisted triple product}
% ===================================================================

¥subsection*{Abstract}

Apply‾$¥sigma$ componentwise to each of the three octonion blocks.
This yields a linear map on‾$¥RR^{24}$, still denoted‾$¥sigma$, that
is an isometry and satisfies $¥sigma^2 = ¥mathrm{id}$.  Replace the
octonion product in each of the nine block-pair products of‾$¥star$
by the $¥sigma$-twisted product $¥cdot_¥sigma$ from
¥eqref{eq:twist}: this produces a new bilinear product
$¥star_¥sigma$ on‾$¥RR^{24}$.  By the same formal manipulation as
before,
¥[
  u ¥star_¥sigma v ¥;=¥; ¥sigma¥bigl(¥sigma(u) ¥star ¥sigma(v)¥bigr),
¥]
where on the right $¥star$ is the standard (untwisted) triple
product.

¥subsection*{Concrete}

Choose $¥sigma = (1¥; 2)$, acting on imaginary indices of each
octonion block.  Then for $u, v ¥in ¥LL$,
¥[
  (¥text{block sum of } u ¥star_¥sigma v)
  ¥;=¥;
  ¥sigma¥bigl((¥sigma(x) + ¥sigma(y) + ¥sigma(z)) ¥cdot
              (¥sigma(x') + ¥sigma(y') + ¥sigma(z'))¥bigr).
¥]
Inside the outer‾$¥sigma$, each triple sum lies in‾$¥sigma(Ls)$
(since each original triple sum lies in‾$Ls$, by Wilson's
condition‾(3) on‾$u, v$).  By
Example‾¥ref{ex:sigmaLs-succeeds}, $¥sigma(Ls) ¥cdot ¥sigma(Ls)
¥subseteq ¥sigma(Ls)$.  Applying the outer‾$¥sigma$ brings the
result into $¥sigma(¥sigma(Ls)) = Ls$.

So the block sum of $u ¥star_¥sigma v$ lies in‾$Ls$: Wilson's
condition‾(3) is restored.  Conditions‾(1) and‾(2) go through by
analogous arguments (see the main paper, Section‾4).  The net effect
is:

¥begin{example}[$¥sigma$-twisted triple product succeeds on‾$¥LL$]
For all $u, v ¥in ¥LL$, the product $u ¥star_¥sigma v$ lies
in‾$¥LL$.  This is Theorem‾3.1 of the main paper.
¥end{example}

¥subsection*{What is twisted and what is invariant (at the Leech
level)}

¥begin{itemize}
¥item ¥textbf{Twisted:} the octonion multiplication used inside
  each of the nine block-pair products.  Equivalently, the Fano
  triples‾¥eqref{eq:fano} are replaced block by block by their
  $¥sigma$-images.
¥item ¥textbf{Invariant under‾$¥sigma$:} the ambient space
  $¥RR^{24}$, the three-block decomposition into‾$¥OO_1, ¥OO_2,
  ¥OO_3$, the $¥ZZ_3$ cross-block routing rule, and the
  sub-lattice‾$L$ of each block ($¥sigma(L) = L$).
¥item ¥textbf{Not invariant under‾$¥sigma$:} the sub-lattice‾$Ls$
  of each block ($¥sigma(Ls) ¥ne Ls$), and through it the Leech
  lattice itself ($¥sigma(¥LL) ¥ne ¥LL$ in general, since‾$¥LL$ is
  defined by a condition involving‾$Ls$).
¥end{itemize}

The comparison with the $E_8$ level is exact.  One level down,
$¥sigma$ fixes‾$L$ but not‾$Ls$, and this is what allows the twist
to change the closure question for‾$Ls$.  One level up, $¥sigma$
fixes‾$¥OO_1 ¥oplus ¥OO_2 ¥oplus ¥OO_3$ but not‾$¥LL$, and this is
what allows the twist to change the closure question for‾$¥LL$.

% ===================================================================
¥section{Could the $¥sigma$-twist undo Coxeter's fix?}
% ===================================================================

A natural concern:¥ Coxeter's correction‾$¥kappa$ of Kirmse swaps two
Fano triples in the multiplication table, producing a convention
under which the $E_8$ lattice‾$L$ closes.  Our‾$¥sigma$ is also a
transposition.  Could a poor choice of‾$¥sigma$ effectively reverse
Coxeter's swap and return us to Kirmse's 1924 regime, where‾$L$
fails to close?

The answer is no, for a reason that is worth surfacing because it
also explains why‾$¥sigma$ can help at all.  Coxeter's‾$¥kappa$ and
our‾$¥sigma$ act at different structural levels:

¥begin{itemize}
¥item Coxeter's‾$¥kappa$ is a transposition of ¥emph{Fano
  triples}.¥ It modifies the octonion multiplication table globally:¥
  the products of ¥emph{all} basis vectors whose Fano lines are
  touched are rewritten.
¥item Our‾$¥sigma$ is a transposition of ¥emph{coordinate axes}.¥ It
  is a linear involution of $¥RR^8$ that extends to the multiplication
  by conjugation, $a ¥cdot_¥sigma b = ¥sigma(¥sigma(a) ¥cdot
  ¥sigma(b))$.
¥end{itemize}

Most coordinate-level transpositions do ¥emph{not} correspond to
Fano-triple permutations; for instance, with our conventions
$¥sigma = (1¥;2)$ does not send the triple $(2,3,5)$ to any of the
other six listed in‾¥eqref{eq:fano}, so the Fano plane's line
structure is not preserved by‾$¥sigma$ at all.¥ In particular,
$¥sigma$ cannot be $¥kappa^{-1}$¥ because‾$¥kappa$ ¥emph{is} a
rearrangement of lines while‾$¥sigma$ is not.

Beyond this structural separation, there is a stronger and more
directly relevant guarantee.  The sub-lattice‾$L$ is invariant
under‾$¥sigma$:
¥[
  ¥sigma(L) = L
  ¥qquad ¥text{(Lemma A of the main paper).}
¥]
Coxeter's correction is a statement about‾$L$:¥ under his corrected
convention, $L ¥cdot L ¥subseteq L$.¥ Because‾$¥sigma$ fixes‾$L$, the
$¥sigma$-twisted product inherits this closure directly:
¥[
  L ¥cdot_¥sigma L
  ¥;=¥; ¥sigma¥bigl(¥sigma(L) ¥cdot ¥sigma(L)¥bigr)
  ¥;=¥; ¥sigma(L ¥cdot L)
  ¥;¥subseteq¥; ¥sigma(L)
  ¥;=¥; L.
¥]
So regardless of which coordinate transposition‾$¥sigma$ we choose
(so long as $¥sigma(L) = L$, which holds for any permutation of the
imaginary axes that preserves the $D_8^+$ structure), Coxeter's $E_8$
closure is automatically preserved.  Kirmse's 1924 regime, which
failed at the level of‾$L$ itself, cannot be reached from Coxeter's
convention by a twist that fixes‾$L$.

What the $¥sigma$-twist ¥emph{does} change is how the product sees
sub-lattices of‾$L$ that‾$¥sigma$ does not fix --- in our case,
$Ls$.¥ Here‾$¥sigma(Ls) ¥ne Ls$ (explicit witness in item‾1 of the
next section), and this is precisely what allows the twist to
resolve Wilson's
condition‾(3) without touching conditions‾(1) or‾(2).  The two
corrections --- Coxeter's‾$¥kappa$ at the Fano-triple level, and
our‾$¥sigma$ at the coordinate level --- are therefore acting on
different strata of the same construction:¥ Coxeter's fixes the
$E_8$ stratum once and for all; ours redirects a strictly
finer-grained sub-lattice question that Coxeter's convention, while
necessary, does not by itself settle.

In short: Coxeter's‾$¥kappa$ is not a linear map on‾$¥RR^8$ at all
(it is a rearrangement of the multiplication table); our‾$¥sigma$
is a linear map on‾$¥RR^8$ that happens to fix‾$L$ on the nose.
There is no route along the $¥sigma$-family back to Kirmse's regime.

% ===================================================================
¥section{Summary of concrete data}
% ===================================================================

¥noindent Three empirical facts are verified by exact rational
arithmetic in
¥href{../python_project/src/symbolic_proof_checks.py}
     {¥texttt{symbolic¥_proof¥_checks.py}}:
¥begin{enumerate}
¥item $¥sigma(Ls) ¥ne Ls$ (explicit witness:
  $(-1, 0, 1, 0, 0, 1, 0, 1) ¥in ¥sigma(Ls) ¥setminus Ls$).
¥item $Ls ¥cdot Ls ¥not¥subseteq Ls$ under the standard octonion
  product (explicit witness‾$a, b$ in
  Example‾¥ref{ex:Ls-fails} with $a ¥cdot b ¥notin Ls$).
¥item $¥sigma(Ls) ¥cdot ¥sigma(Ls) ¥subseteq ¥sigma(Ls)$ under the
  standard octonion product (all 64 basis products land in
  $¥sigma(Ls)$).
¥end{enumerate}

¥noindent Two equivalent phrasings of the main result are:
¥begin{itemize}
¥item ``$¥LL$ is closed under the $¥sigma$-twisted triple product
  $¥star_¥sigma$.''
¥item ``$¥sigma(¥LL)$ is closed under the standard triple product
  $¥star$.''
¥end{itemize}
Both say the same thing via the algebra isomorphism
$¥sigma: (¥RR^{24}, ¥star) ¥to (¥RR^{24}, ¥star_¥sigma)$; neither
says that $¥LL$ is closed under‾$¥star$, which is false.

% ===================================================================

¥end{document}
