\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}

% ---------------------------------------------------------------
% Theorem environments
% ---------------------------------------------------------------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ---------------------------------------------------------------
% Notation
% ---------------------------------------------------------------
\newcommand{\OO}{\mathbb{O}}          % octonions
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\HH}{\mathbb{H}}
\newcommand{\LL}{\Lambda}             % Leech lattice
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\Nrm}{N}                  % norm form
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Co}{Co}

\title[An order on the Leech lattice]{An order on the Leech lattice\\
from transposition-twisted octonions}

\author{Jens Koeplinger}

\address{105 E Avondale Dr, Greensboro, NC 27403, USA}
\email{jenskoeplinger@gmail.com}

\date{14 April 2026}

\subjclass[2020]{11H06, 17A75, 11H56}
\keywords{Leech lattice, octonions, non-associative order, integral algebra}

\begin{document}

\begin{abstract}
We show that the Leech lattice~$\LL$ admits the structure of an order
in a 24-dimensional real algebra.  The product is built from three
identical copies of a \emph{transposition-twisted} octonion algebra ---
the algebra obtained by applying a single transposition to the
imaginary basis elements of the standard Fano-plane multiplication ---
with $\ZZ_3$-symmetric cross-block routing.  The construction sits at
the end of a short chain of transpositions: Kirmse
$\to$ Coxeter $\to$ $\sigma$-twist, each step a transposition of two
basis indices that resolves a closure obstruction at the next level
--- Coxeter's transposition turns Kirmse's~$E_8$ candidate into an
order in the octonions, and the present $\sigma$-twist turns the
triple-octonion product on~$\LL$ into an order.

This result was developed through a systematic human--AI collaboration;
the complete research record is publicly available~\cite{repo}.
\end{abstract}

\maketitle

% ===============================================================
\section{Introduction}\label{sec:intro}
% ===============================================================

The Leech lattice~$\LL$ is the unique even unimodular lattice of
rank~24 with no vectors of squared norm~2
\cite{ConwaySloane1999}.  Its minimal shell $\operatorname{Min}(\LL)$
consists of 196{,}560 vectors of squared norm~8, and its automorphism
group is the Conway group~$\Co_0$, a double cover of the sporadic
simple group~$\Co_1$.

In rank~8, the $E_8$ lattice~$L$ carries an octonionic multiplication
that makes it a maximal order in the real octonion division algebra.
There are seven such orders, one for each of the seven maximal orders
of the integral octonions \cite{Coxeter1946, ConwaySmith2003}.  Wilson
\cite{Wilson2009} showed that $\LL$ embeds naturally in~$L^3$ via three
sublattice conditions, raising the question: does $\LL$ inherit an
algebraic structure from the octonion product?

This question has been studied from several angles.  Dixon
\cite{Dixon2010} constructed~$\LL$ using an octonionic XY-product.
Baez and Egan \cite{Baez2014} explored Jordan-algebraic structures
on~$\LL$ within the exceptional Jordan algebra~$\mathfrak{h}_3(\OO)$,
finding closure under a doubled Jordan product.

The standard triple-octonion product on~$L^3$ --- using the same
octonion multiplication in all three blocks with
$\ZZ_3$-symmetric cross-block routing --- satisfies two of Wilson's
three sublattice conditions automatically, but closure on the
third requires a block-sum sublattice~$Ls$ to be preserved under
octonion multiplication, and the standard product does not
preserve it.

The main result of this paper is that a single transposition applied to
the Fano-plane triples properly closes the remaining condition.
The key insight is that the twist maps the block-sum sublattice~$Ls$
to~$\sigma(Ls)$, a different sublattice of~$L$ that \emph{is} closed
under the standard octonion product.

\begin{theorem}\label{thm:main}
Let\/ $\cdot_\sigma$ denote the transposition-twisted octonion
multiplication on~$\RR^8$ (Definition~\ref{def:twist}), and let\/
$\star$ denote the associated triple product on
$\RR^{24} = \OO_1 \oplus \OO_2 \oplus \OO_3$ with $\ZZ_3$ cross-block
routing (Definition~\ref{def:triple}).  Then for all
$u, v \in \LL$, the product $u \star v$ lies in~$\LL$.
\end{theorem}

\begin{corollary}\label{cor:order}
$(\LL, +, \star)$ is an order in the $\RR$-algebra
$(\RR^{24}, +, \star)$.
\end{corollary}

\begin{proof}
$\operatorname{Min}(\LL)$ generates~$\LL$ over~$\ZZ$
\cite{ConwaySloane1999}.  The product~$\star$ is bilinear (it is a
sum of bilinear octonion products).  By Theorem~\ref{thm:main},
$\star$ maps $\LL \times \LL$ into~$\LL$.
\end{proof}

\begin{remark}[Order-closure, not shell-closure]\label{rem:order-closure}
The closure statement $\LL \star \LL \subseteq \LL$ means that
products land somewhere in~$\LL$; it does not assert that the
minimal shell is preserved.  Empirically, the product norm on
$\operatorname{Min}(\LL) \times \operatorname{Min}(\LL)$ takes values
in $\{16, 32, 48, 64, 80, 96, 112, 128\}$ (see
Section~\ref{sec:properties}), none of which equals the minimal value
$\Nrm = 8$: products of minimal vectors always land on strictly
higher shells.  This is consistent with $(\LL, +, \star)$ being an
order in the ring-theoretic sense --- a subring that is a free
$\ZZ$-module of rank equal to the ambient algebra's dimension ---
a property weaker than, and independent of, any shell-level closure.
\end{remark}

% ===============================================================
\section{Preliminaries}\label{sec:prelim}
% ===============================================================

\subsection{The octonion algebra}

Let $\OO$ denote the real octonion algebra with basis
$\{e_0, e_1, \ldots, e_7\}$, where $e_0$ is the identity.
The multiplication is determined by the seven \emph{Fano triples}
\begin{equation}\label{eq:fano}
  (1,2,4),\; (2,3,5),\; (3,4,6),\; (4,5,7),\; (5,6,1),\;
  (6,7,2),\; (7,1,3),
\end{equation}
encoding the rule: for each triple $(a,b,c)$,
$e_a \cdot e_b = +e_c$, with cyclic permutations carrying a positive
sign and anti-cyclic permutations carrying a negative sign
\cite{Baez2002}.  Every imaginary unit squares to $-e_0$.  The
algebra~$\OO$ is the unique real normed division algebra of
dimension~8 \cite{Baez2002}.

The \emph{norm form} is $\Nrm(x) = x \bar{x} = \sum_{i=0}^7 x_i^2$
for $x = \sum_i x_i e_i$, where $\bar{x} = x_0 e_0 - \sum_{i=1}^7
x_i e_i$ is the conjugate.  The composition property
$\Nrm(xy) = \Nrm(x)\Nrm(y)$ holds \cite{Baez2002}.

\subsection{The $E_8$ lattice}

Wilson's $E_8$ lattice is $L = D_8^+$, the set of vectors in~$\RR^8$
whose coordinates (in the basis $\{e_0, \ldots, e_7\}$) are either all
integers or all half-integers, with even coordinate sum
\cite{Wilson2009, ConwaySmith2003}.  It has 240~roots (vectors of
squared norm~2).  Under the standard octonion multiplication,
$L$~is a maximal order in~$\OO$: it is closed under multiplication
\cite{Coxeter1946, ConwaySmith2003}.

\begin{remark}[Kirmse's convention versus ours]\label{rem:kirmse}
The closure $L \cdot L \subseteq L$ depends on which Fano-plane
labelling defines the octonion product.  Kirmse~\cite{Kirmse1924}
originally proposed a specific labelling under which $D_8^+$ is
\emph{not} closed --- a subtle error corrected by
Coxeter~\cite{Coxeter1946}, who showed that permuting two basis
indices yields a different multiplication on the same~$\RR^8$ under
which $D_8^+$ \emph{is} closed.  There are seven such
Coxeter-corrected conventions, each giving one of the seven maximal
orders of the integral octonions~\cite{ConwaySmith2003}; see
also Petersson~\cite{Petersson2018} for a modern account.

The ``standard'' Fano triples~\eqref{eq:fano} used in this paper are
one of the Coxeter-corrected conventions, not Kirmse's.  We have
verified $L \cdot L \subseteq L$ directly under our convention: using
an explicit $\ZZ$-basis of $L$ and exact rational arithmetic, all 64
basis products are integer combinations of the basis.
\end{remark}

\begin{remark}[Closure of the lattice, not the shells]\label{rem:shells}
When we say $L$ is closed under multiplication, we mean closure of the
lattice~$L$, not of any individual shell.  Since the octonion product
is norm-multiplicative, the product of two roots (squared norm~2) has
squared norm~4 and lies on the second shell of~$L$, not among the
roots.

The same distinction will apply to $(\LL, +, \star)$ in
Theorem~\ref{thm:main}: products of minimal Leech vectors land on
higher shells of~$\LL$, not back in the minimal shell (see
Section~\ref{sec:properties}).
\end{remark}

\subsection{Wilson's Leech lattice}\label{subsec:wilson-leech}

Define $s = \tfrac{1}{2}(-e_0 + e_1 + \cdots + e_7)$ and
$\bar{s} = \tfrac{1}{2}(-e_0 - e_1 - \cdots - e_7)$.  Both lie in~$L$
with $\Nrm(s) = \Nrm(\bar{s}) = 2$.  Let
\[
  L\bar{s} = \{ \ell \cdot \bar{s} : \ell \in L \}, \qquad
  Ls = \{ \ell \cdot s : \ell \in L \}.
\]
These are sublattices of~$L$ with $[L : L\bar{s}] = [L : Ls] = 16$,
$L\bar{s} + Ls = L$, and $L\bar{s} \cap Ls = 2L$
\cite{Wilson2009}.

\begin{definition}[Wilson {\cite[Section~3]{Wilson2009}}]
\label{def:wilson-leech}
The \emph{Leech lattice} is
\[
  \LL = \bigl\{ (x,y,z) \in L^3 :
    \text{(1) } x,y,z \in L;\;
    \text{(2) } x{+}y,\, x{+}z,\, y{+}z \in L\bar{s};\;
    \text{(3) } x{+}y{+}z \in Ls
  \bigr\}.
\]
\end{definition}

The 196{,}560 minimal vectors of~$\LL$ decompose into three families
\cite{Wilson2009}:

\smallskip
\begin{tabular}{@{}lll@{}}
\toprule
Family & Formula & Count \\
\midrule
Type~1 & $(2\lambda, 0, 0)$ and cyclic permutations & $3 \times 240 = 720$ \\
Type~2 & $(\lambda\bar{s},\; (\lambda\bar{s})j,\; 0)$ and cyclic
  perms & $3 \times 240 \times 16 = 11{,}520$ \\
Type~3 & $((\lambda s)j,\; \lambda k,\; (\lambda j)k)$ and cyclic
  perms & $3 \times 240 \times 16^2 = 184{,}320$ \\
\bottomrule
\end{tabular}
\smallskip

\noindent
Here $\lambda$ ranges over the 240~roots of~$L$, and
$j, k \in J = \{\pm e_t : 0 \le t \le 7\}$ (16~unit octonions).

% ===============================================================
\section{The construction}\label{sec:construction}
% ===============================================================

\begin{definition}[Transposition-twisted octonion algebra]
\label{def:twist}
Let $\sigma = (s\; t)$ be a transposition on $\{1, \ldots, 7\}$.
Extend~$\sigma$ to a linear map $\sigma: \RR^8 \to \RR^8$ that fixes
$e_0$ and permutes $e_s \leftrightarrow e_t$.  Apply~$\sigma$ to each
standard Fano triple: $(a,b,c) \mapsto (\sigma(a), \sigma(b), \sigma(c))$.
The resulting triples define a new multiplication table on~$\RR^8$.
Equivalently, the \emph{$\sigma$-twisted multiplication} is
\begin{equation}\label{eq:twist-formula}
  x \cdot_\sigma y \;=\; \sigma\bigl(\sigma(x) \cdot \sigma(y)\bigr),
\end{equation}
where $\cdot$ denotes the standard octonion product.
\end{definition}

\begin{proposition}\label{prop:iso}
$(\RR^8, \cdot_\sigma)$ is an octonion algebra, isomorphic
to~$(\OO, \cdot)$ via the map~$\sigma$ itself:
\begin{equation}\label{eq:iso}
  \sigma(x \cdot y) = \sigma(x) \cdot_\sigma \sigma(y)
  \qquad \text{for all } x, y \in \RR^8.
\end{equation}
In particular, $\cdot_\sigma$ satisfies the composition property
$\Nrm(x \cdot_\sigma y) = \Nrm(x)\Nrm(y)$.
\end{proposition}

\begin{proof}
For any $x, y \in \RR^8$:
\[
  \sigma(x) \cdot_\sigma \sigma(y)
  = \sigma\bigl(\sigma(\sigma(x)) \cdot \sigma(\sigma(y))\bigr)
  = \sigma(x \cdot y),
\]
since $\sigma^2 = \mathrm{id}$.  Thus $\sigma: (\OO, \cdot) \to
(\RR^8, \cdot_\sigma)$ is a vector-space isomorphism that preserves
multiplication.  Composition follows.
\end{proof}

\begin{remark}\label{rem:table-diffs}
For the transposition $\sigma = (1\;2)$, exactly 30 of the 64
multiplication table entries change relative to the standard table:
6~entries undergo a sign change only, 16~undergo a target-index change
only, and 8~undergo both.
\end{remark}

\begin{remark}\label{rem:unique}
The 21~transpositions of $\{1, \ldots, 7\}$ form a single orbit under
the Fano-plane automorphism group $\GL(3, \FF_2)$ (which has order~168
and acts on the 7~points of the Fano plane).  Therefore all
transposition twists produce the same multiplication table up to basis
relabelling.  The construction is essentially unique.
\end{remark}

\begin{definition}[Triple product with $\ZZ_3$ routing]
\label{def:triple}
Decompose $\RR^{24} = \OO_1 \oplus \OO_2 \oplus \OO_3$ into three
copies of~$\RR^8$ (blocks indexed $\alpha = 1, 2, 3$).  For
$u = (x, y, z)$ and $v = (x', y', z')$ with $x, x' \in \OO_1$, etc.,
define the \emph{triple product}
\begin{equation}\label{eq:triple}
  u \star v = (P_1,\, P_2,\, P_3),
\end{equation}
where each block $P_\alpha$ is the sum of three $\sigma$-twisted
octonion products:
\begin{align}
  P_1 &= x \cdot_\sigma x' + z \cdot_\sigma y' + y \cdot_\sigma z',
  \label{eq:P1} \\
  P_2 &= y \cdot_\sigma y' + x \cdot_\sigma z' + z \cdot_\sigma x',
  \label{eq:P2} \\
  P_3 &= z \cdot_\sigma z' + y \cdot_\sigma x' + x \cdot_\sigma y'.
  \label{eq:P3}
\end{align}
The routing rule is: a product of blocks $\alpha$ and~$\beta$ lands in
block~$\gamma$, where $\{\alpha, \beta, \gamma\} = \{1,2,3\}$ for
$\alpha \ne \beta$, and in block~$\alpha$ when $\alpha = \beta$.
All nine block-pair products use the same multiplication~$\cdot_\sigma$.
\end{definition}

% ===============================================================
\section{Proof of closure}\label{sec:proof}
% ===============================================================

We prove Theorem~\ref{thm:main} by showing that the twisted product
preserves each of Wilson's three conditions
(Definition~\ref{def:wilson-leech}).  The proof rests on four
lemmas about the interaction between the transposition~$\sigma$, the
maximal order~$L$, and the sublattices $Ls$ and~$L\bar{s}$.

\subsection{Reduction to standard products}

Since each $\sigma$-twisted product has the form
$a \cdot_\sigma b = \sigma(\sigma(a) \cdot \sigma(b))$, and $\sigma$
is linear, each block $P_\alpha$ is $\sigma$ applied to a sum of
standard octonion products of $\sigma$-images of the inputs.

Write $X = \sigma(x)$, $Y = \sigma(y)$, $Z = \sigma(z)$,
$X' = \sigma(x')$, $Y' = \sigma(y')$, $Z' = \sigma(z')$.  Then:
\begin{align}
  P_1 &= \sigma\bigl(X \cdot X' + Z \cdot Y' + Y \cdot Z'\bigr),
  \label{eq:P1-std} \\
  P_2 &= \sigma\bigl(Y \cdot Y' + X \cdot Z' + Z \cdot X'\bigr),
  \label{eq:P2-std} \\
  P_3 &= \sigma\bigl(Z \cdot Z' + Y \cdot X' + X \cdot Y'\bigr).
  \label{eq:P3-std}
\end{align}

Taking sums:
\begin{itemize}
\item \textbf{Block sum.}
  $P_1 + P_2 + P_3 = \sigma\bigl((X+Y+Z) \cdot (X'+Y'+Z')\bigr)$.
\item \textbf{Pairwise sums.}  For example,
  $P_1 + P_2 = \sigma\bigl(X \cdot (X'+Z') + Y \cdot (Y'+Z') +
  Z \cdot (X'+Y')\bigr)$, and similarly for the other two pairs.
\end{itemize}

\subsection{The four lemmas}

\begin{lemma}\label{lem:sigma-L}
$\sigma(L) = L$.
\end{lemma}

\begin{proof}
The lattice $L = D_8^+$ consists of all vectors in~$\RR^8$ whose
coordinates are either all integers or all half-integers, with even
coordinate sum.  The transposition~$\sigma$ permutes two coordinates,
which preserves both the integer/half-integer parity and the coordinate
sum.  Hence $\sigma(L) \subseteq L$.  Since $\sigma$ is an involution,
$\sigma(L) = L$.
\end{proof}

\begin{lemma}\label{lem:L-order}
$L \cdot L \subseteq L$.
\end{lemma}

\begin{proof}
$L$ is a maximal order in~$\OO$
\cite{Coxeter1946, ConwaySmith2003}.
\end{proof}

\begin{lemma}\label{lem:L-sigmaLsbar}
$L \cdot \sigma(L\bar{s}) \subseteq \sigma(L\bar{s})$.
\end{lemma}

\begin{proof}
Both $L$ and $\sigma(L\bar{s})$ are lattices of rank~8.  It suffices to
check the inclusion on $\ZZ$-bases: if $\{b_1, \ldots, b_8\}$ is a
$\ZZ$-basis of~$L$ and $\{c_1, \ldots, c_8\}$ is a $\ZZ$-basis
of~$\sigma(L\bar{s})$, then $L \cdot \sigma(L\bar{s}) \subseteq
\sigma(L\bar{s})$ holds if and only if each of the 64 products
$b_i \cdot c_j$ is an integer combination of $\{c_1, \ldots, c_8\}$.
This is verified by exact arithmetic (using rational numbers, no
floating point) on explicit bases.  All 64 products lie
in~$\sigma(L\bar{s})$.
\end{proof}

\begin{lemma}\label{lem:sigmaLs-closed}
$\sigma(Ls) \cdot \sigma(Ls) \subseteq \sigma(Ls)$.
\end{lemma}

\begin{proof}
As in Lemma~\ref{lem:L-sigmaLsbar}, verified on $\ZZ$-bases by exact arithmetic: all 64
products of basis vectors of~$\sigma(Ls)$ lie in~$\sigma(Ls)$.
\end{proof}

\begin{remark}\label{rem:Ls-not-closed}
The sublattice~$Ls$ is \emph{not} closed under the standard octonion
product: there exist basis vectors $a, b \in Ls$ with $a \cdot b
\notin Ls$.  This is precisely the obstruction that the twist
resolves: it moves the block-sum sublattice from~$Ls$ to~$\sigma(Ls)$,
where closure holds by Lemma~\ref{lem:sigmaLs-closed}.

Note that $\sigma(Ls)$ is genuinely different from~$Ls$ (not just a
relabelling).  For example, with $\sigma = (1\;2)$, the element
$v = (-1, 0, 1, 0, 0, 1, 0, 1)$ lies in~$\sigma(Ls)$ but direct
computation shows it is not an integer combination of any $\ZZ$-basis
for~$Ls$, so $v \notin Ls$.  This distinction is what makes the twist
useful: $\sigma$ is not the identity on the block-sum sublattice, so
it can redirect a product into a different (and, as it turns out,
closed) sublattice.
\end{remark}

\subsection{Condition~(1): each block lies in~$L$}

\begin{proposition}\label{prop:cond1}
For all $u, v \in \LL$, each block $P_\alpha$ of $u \star v$ lies
in~$L$.
\end{proposition}

\begin{proof}
Consider $P_1 = \sigma(X \cdot X' + Z \cdot Y' + Y \cdot Z')$.
Since $u \in \LL$, condition~(1) gives $x \in L$, so
$X = \sigma(x) \in L$ by Lemma~\ref{lem:sigma-L}.  Similarly
$X', Y, Y', Z, Z' \in L$.
By Lemma~\ref{lem:L-order}, each product $X \cdot X'$, $Z \cdot Y'$,
$Y \cdot Z'$ lies in~$L$.  Their sum lies in~$L$ (closed under
addition).  Applying $\sigma$ and using Lemma~\ref{lem:sigma-L},
$P_1 \in L$.  The same argument applies to
$P_2$ and~$P_3$.
\end{proof}

\subsection{Condition~(2): pairwise sums lie in~$L\bar{s}$}

\begin{proposition}\label{prop:cond2}
For all $u, v \in \LL$, the pairwise sums $P_\alpha + P_\beta$ lie
in~$L\bar{s}$.
\end{proposition}

\begin{proof}
We show $P_1 + P_2 \in L\bar{s}$; the other two cases follow by the
$\ZZ_3$ symmetry of the routing.

From \eqref{eq:P1-std}--\eqref{eq:P2-std}:
\[
  P_1 + P_2 = \sigma\bigl(X \cdot (X' + Z') + Y \cdot (Y' + Z')
    + Z \cdot (X' + Y')\bigr).
\]

Since $v \in \LL$, condition~(2) gives $x' + z' \in L\bar{s}$, so
$X' + Z' = \sigma(x' + z') \in \sigma(L\bar{s})$.  Similarly
$Y' + Z' \in \sigma(L\bar{s})$ and $X' + Y' \in \sigma(L\bar{s})$.
Since $u \in \LL$, we have $X, Y, Z \in L$ (by Lemma~\ref{lem:sigma-L}).

By Lemma~\ref{lem:L-sigmaLsbar}, each product $X \cdot (X'+Z')$, $Y \cdot (Y'+Z')$,
$Z \cdot (X'+Y')$ lies in~$\sigma(L\bar{s})$.  Their sum lies
in~$\sigma(L\bar{s})$.  Applying the outer~$\sigma$:
\[
  P_1 + P_2 = \sigma(\text{element of } \sigma(L\bar{s}))
  \in \sigma(\sigma(L\bar{s})) = L\bar{s},
\]
since $\sigma^2 = \mathrm{id}$.
\end{proof}

\subsection{Condition~(3): the block sum lies in~$Ls$}

This is the crux of the proof.  For the standard (untwisted) product
the block sum would require $Ls \cdot Ls \subseteq Ls$, which does
not hold (Remark~\ref{rem:Ls-not-closed}).  The twist redirects the
computation through~$\sigma(Ls)$, where closure holds.

\begin{proposition}\label{prop:cond3}
For all $u, v \in \LL$, the block sum $P_1 + P_2 + P_3$ lies in~$Ls$.
\end{proposition}

\begin{proof}
From \eqref{eq:P1-std}--\eqref{eq:P3-std}:
\[
  P_1 + P_2 + P_3 = \sigma\bigl((X+Y+Z) \cdot (X'+Y'+Z')\bigr).
\]

Since $u \in \LL$, condition~(3) gives $x + y + z \in Ls$, so
\[
  X + Y + Z = \sigma(x+y+z) \in \sigma(Ls).
\]
Similarly $X' + Y' + Z' \in \sigma(Ls)$.

By Lemma~\ref{lem:sigmaLs-closed}, $\sigma(Ls) \cdot \sigma(Ls) \subseteq \sigma(Ls)$, so
\[
  (X+Y+Z) \cdot (X'+Y'+Z') \in \sigma(Ls).
\]
Applying the outer~$\sigma$:
\[
  P_1 + P_2 + P_3 \in \sigma(\sigma(Ls)) = Ls. \qedhere
\]
\end{proof}

This completes the proof of Theorem~\ref{thm:main}.
% ===============================================================
\section{Algebraic properties}\label{sec:properties}
% ===============================================================

The order $(\LL, +, \star)$ has been tested for several standard
algebraic properties.  The results are summarised below.

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
Property & Holds? & Details \\
\midrule
Multiplicative identity & No & $(e_0, e_0, e_0)$ has $\Nrm = 3$;
  not in $\LL$ \\
Commutativity & No & ${<}\,0.1\%$ of pairs commute \\
Norm multiplicativity & No & $\approx 47\%$ of pairs satisfy
  $\Nrm(u \star v) = \Nrm(u)\Nrm(v)$ \\
Left alternativity & No & ${<}\,0.1\%$ of pairs satisfy
  $(x \star x) \star y = x \star (x \star y)$ \\
Right alternativity & No & ${<}\,0.1\%$ of pairs satisfy
  $(x \star y) \star y = x \star (y \star y)$ \\
Flexibility & No & $\approx 8.6\%$ of pairs satisfy
  $(x \star y) \star x = x \star (y \star x)$ \\
Power-associativity & No & $\approx 29\%$ of elements satisfy
  $x^2 \star x = x \star x^2$ \\
\bottomrule
\end{tabular}
\end{center}

\begin{remark}
The product norm on $\operatorname{Min}(\LL) \times
\operatorname{Min}(\LL)$ takes values in
$\{16, 32, 48, 64, 80, 96, 112, 128\}$, all multiples of~16.
The mode is~64 (the norm-multiplicative value
$\Nrm(u)\Nrm(v) = 8 \times 8$).
\end{remark}

\begin{remark}
The order is non-unital: the ambient algebra $(\RR^{24}, +, \star)$
has multiplicative identity $(e_0, e_0, e_0)$, but this element has
squared norm~3 and does not lie in~$\LL$.  Thus $(\LL, +, \star)$ is
a non-unital order.
\end{remark}

The order does not satisfy any of the classical algebraic identities
(associativity, alternativity, flexibility, power-associativity,
commutativity).  The defining property is \emph{closure}: the Leech
lattice is stable under the product.  The relationship between
$\star$ and the Conway group~$\Co_0$ (the automorphism group of~$\LL$
as a lattice) is an open question.

% ===============================================================
\section{Research methodology}\label{sec:methodology}
% ===============================================================

This research was conducted as a collaboration between a human
researcher and an AI agent (Claude, Anthropic) over a series of
directed prompts, governed by a written methodology (the
``Manifesto'') established at the outset.

The human researcher (the author) brought the initial hypothesis ---
motivated by prior work on Okubo algebras and autotopies of normed
composition algebras --- established the research methodology, and
maintained control of the inquiry through prompt-by-prompt review.
The AI agent performed the computational work: implementing octonion
algebras, Wilson's Leech lattice construction, trial algebras, and
closure tests, and contributed mathematical analysis.

The computational verification was a vehicle in the discovery process,
designed to mediate distrust into received knowledge and intermediate
results.  Tests were written from first principles to independently
verify every foundation: that octonion multiplication satisfies the
composition property $\Nrm(xy) = \Nrm(x)\Nrm(y)$, that Wilson's
construction produces a lattice with the correct minimal shell of
196{,}560 vectors, and that the $E_8$ lattice is indeed a maximal
order.  This independent verification is how we know that the
properties stated in the literature are correctly implemented.
Closure of the twisted product was confirmed on over 12{,}000{,}000
random pairs of minimal vectors with zero failures before the
symbolic proof was found.

The symbolic proof (Section~\ref{sec:proof}) and the exact-arithmetic
verifications of Lemmas~\ref{lem:L-sigmaLsbar} and~\ref{lem:sigmaLs-closed} use Python's \texttt{fractions.Fraction}
type, operating entirely in~$\mathbb{Q}$ with no floating-point
arithmetic.  The $\ZZ$-basis checks involve solving 64 systems of
8~linear equations each, all yielding integer solutions.

The methodology required:
\begin{itemize}
\item Every claim backed by a reference or computation.
\item Every prompt logged, with actions recorded.
\item Every conclusion independently verified.
\item Forward-evolving corrections: no rewriting history; errors are
  corrected in subsequent entries.
\end{itemize}

The path to the result was not direct.  Six trial algebra families
were explored and ruled out before the transposition twist was
identified in trial~007.  The systematic methodology ---
understanding the anatomy of failure (which conditions fail, for which
vector types, and why) before trying the next approach --- was
essential.  The failure analysis of the untwisted product (Wilson's
condition~(3) per Definition~\ref{def:wilson-leech}, exclusively
type-3$\times$type-3) directly pointed to the
mechanism that the twist needed to correct.

Iterative convergence toward a robust result --- including learning
along the way what it is we are actually doing --- is how study,
research, and discovery have always proceeded.  The AI agent violated
its own logging protocol on three occasions (prompts 011, 018--019,
and 029--031), each time caught by the human researcher on review.
Conversely, the author learned from the AI agent about errors and
misconceptions in the framing of certain prompts, which likewise
are part of the record.  What is novel here is the ability to trace
the entire path, beginning to end, through the complete list of
prompts.

The full record --- all prompts, code, evidence, and git history ---
is available at the repository~\cite{repo}.  The reader can form
their own assessment of what this means for questions of discovery
and attribution.

% ===============================================================
\section{Related work}\label{sec:related}
% ===============================================================

\textbf{Wilson (2009)} \cite{Wilson2009} characterises~$\LL$ as
$\{(x,y,z) \in L^3 : \text{conditions 1--3}\}$, gives an explicit
listing of the 196{,}560 minimal vectors in the form
$3 \times 240 \times (1 + 16 + 16 \times 16)$, and proves that~$\LL$
is even and self-dual.  The sublattice framework is
the setting used in the present paper.

\textbf{Dixon (2010)} \cite{Dixon2010} constructs~$\LL$ using the
octonion XY-product
\[
  A \circ_{XY} B \;=\; (A X)(Y^{\dagger} B),
\]
where $X, Y \in \OO$ are chosen so that $Y X^{\dagger}$ is a unit
octonion (the identity of the new multiplication).  When $Y^{\dagger} = X^{-1}$
the XY-product specialises to the X-product $A \circ_X B = (AX)(X^{-1}
B)$; Dixon uses this X-product with $X = \ell_0 = \frac{1}{2}(1 + e_1
+ \cdots + e_7)$ to obtain~$E_8$, and general XY-products for
higher-dimensional laminated lattices including~$\LL$.

\textbf{Baez and Egan (2014)} \cite{Baez2014} study closure
properties of integer sublattices of the exceptional Jordan algebra
$\mathfrak{h}_3(\OO)$.  In parts~9--10 of the series, Baez considers
the 27-dimensional lattice $\ZZ^3 \oplus L_L \subset \mathfrak{h}_3(\OO)$,
with integers on the diagonal and an off-diagonal Leech sublattice
$L_L \cong \LL$ embedded as an octonionic triple.  This lattice is
\emph{not} closed under the Jordan product $x \circ y = xy + yx$,
but \emph{is} closed under the doubled product $2(xy + yx)$.
Egan further enumerated at least $17{,}280$ distinct off-diagonal
Leech embeddings $L_L \subset \OO^3$ for which the associated
27-dimensional lattice is closed under this doubled product.
Baez left open the question whether any such lattice is closed under
the undoubled Jordan product --- which would remove the factor-of-2
discrepancy --- and conjectures that the answer is negative.

\textbf{Kirmse (1924)} \cite{Kirmse1924} exhibited the $E_8$ lattice
as a unimodular integral quadratic lattice inside the real octonion
algebra.  His lattice is not closed under multiplication, an error
corrected by Coxeter~\cite{Coxeter1946} by swapping two basis indices
in the Fano-plane labelling.

The present construction continues the same conceptual line.  The
Kirmse--Coxeter step modifies the octonion multiplication (a single
transposition of basis indices) so that the $E_8$ lattice becomes an
order.  The present paper applies a second transposition of basis
indices --- the twist $\sigma$ in Definition~\ref{def:twist} --- to
the already-corrected octonion product, so that the triple-product
construction on the Leech lattice becomes an order.  The chain is
Kirmse $\to$ Coxeter $\to$ $\sigma$-twist: each step is a
transposition, and each resolves a closure obstruction at the next
level of the construction.

% ===============================================================
\section{Conclusion}\label{sec:conclusion}
% ===============================================================

The Leech lattice admits an order under a remarkably simple
construction: three identical copies of a transposition-twisted
octonion algebra with $\ZZ_3$ routing.  The symbolic proof
(Section~\ref{sec:proof}) reduces closure to four lemmas, of which
the crux is Lemma~\ref{lem:sigmaLs-closed}: the sublattice $\sigma(Ls)$ --- unlike~$Ls$
itself --- is closed under the standard octonion product.  The twist
maps the condition-3 sublattice to this closed sublattice, and the
involutory nature of~$\sigma$ maps the result back.

Several questions remain immediately open:
\begin{itemize}
\item The automorphism group of $(\LL, +, \star)$ and its relationship
  to the Conway group~$\Co_0$.
\item Whether the order is maximal in any appropriate sense.
\item Classification of the algebra $(\RR^{24}, +, \star)$: it is
  non-associative, non-alternative, non-flexible, and non-unital
  (as an order), placing it outside the classical families.
\item Whether other permutations of the Fano triples (beyond
  transpositions) also yield closed products on~$\LL$.
\end{itemize}

% ===============================================================
\section{Outlook}\label{sec:outlook}
% ===============================================================

The intention in publishing this result at this stage is to invite
research rather than to sit on the ideas.  If others arrive at deeper
answers faster, so much the better.  What follows is an outlook that
names the directions we find most promising and the sources that
shaped the intuition leading to the present construction.

\subsection{A ternary reformulation}

A more natural classification of $(\RR^{24}, +, \star)$ may emerge
from a \emph{ternary} rather than binary viewpoint.  The Leech lattice
possesses a prominent $\ZZ_3$-symmetry (the cyclic permutation of the
three octonion blocks used throughout this paper), and the binary
product~$\star$ is assembled from nine bilinear building blocks whose
routing is itself governed by that $\ZZ_3$-action.  Ternary
composition algebras, as studied by
Elduque~\cite{Elduque1996,Elduque2000_Triality,Elduque2023_IsotropicNorm},
are typically non-associative, non-commutative, non-alternative, and
non-power-associative --- the same negative classification our order
satisfies --- yet carry high internal symmetry, with order-3
automorphisms playing the central role: the Okubo algebra has
automorphism group $\mathrm{SU}(3)$, and the Okubo product itself
is built from an order-3 automorphism of the octonions.
Complementary perspectives on the Okubo algebra and triality include
the quasigroup-theoretic treatment of Smith and
Vojt\v{e}chovsk\'y~\cite{SmithVojtechovsky2022}, Kamiya and Okubo's
study of algebras satisfying triality and $S_4$-symmetry
\cite{KamiyaOkubo2015}, and Okubo's seminal
book~\cite{Okubo1995_Book}.  A deeper precursor
is Hall's monograph~\cite{Hall2019_Moufang}, which establishes that
Moufang loops and groups with triality are essentially the same
object, giving a structural scaffolding for any triality-based
reformulation.  Recasting $\star$ as a trace-like reduction of a
ternary operation compatible with the $\ZZ_3$-block symmetry is a
natural direction for further work.

\subsection{Consequences for algebraic approaches to physics}

A classification of the ambient algebra $(\RR^{24}, +, \star)$
matters beyond pure algebra.  Programmes that seek to model
fundamental physics on algebraic structures --- especially those that
use the four normed division algebras, triality, and related
symmetric composition algebras --- have grown considerably in recent
years.  The author previously invited~\cite{Koeplinger2023} for
research into autotopies of normed composition algebras as a
mathematical setting for this line of inquiry.  Recent work of Furey and
Hughes~\cite{FureyHughes2025_TrioOfTrialities} isolates triality
itself as the operative structural skeleton: the Standard Model's
$\mathfrak{su}(3) \oplus \mathfrak{su}(2) \oplus \mathfrak{u}(1)$
gauge symmetries are embedded inside
$\mathfrak{tri}(\CC) \oplus \mathfrak{tri}(\HH) \oplus \mathfrak{tri}(\OO)$,
and the triality skeleton is shown to accommodate a variety of
particle-family representations --- the authors' own assignment of
three Standard Model generations to a triality triple in
$\CC \otimes \HH \otimes \OO$ is one concrete realization of that
wider scaffolding.

\subsection{On the form of this work}

This paper is the outcome of a human--AI collaboration conducted
entirely through a sequence of AI prompts (Section~\ref{sec:methodology}),
with the complete record of prompts, code, intermediate results, and
corrections preserved verbatim~\cite{repo}.  The form itself is part
of the contribution.  Collaborations of this kind will become more
common, and this work is offered in its present form partly in the
hope that the process --- how a research question was framed, how the
AI was directed and corrected, how false trails were recognised and
closed --- is instructive for mathematicians considering AI-assisted
research.  Comments on both substance and method are welcome.

% ===============================================================
\section*{Acknowledgments}
% ===============================================================

The author thanks Geoffrey Dixon for helpful discussions and advice
on octonionic structures associated with the Leech lattice.

% ===============================================================
% References
% ===============================================================

\begin{thebibliography}{99}

\bibitem{Baez2002}
J.\,C.~Baez,
\emph{The octonions},
Bull.\ Amer.\ Math.\ Soc.\ \textbf{39} (2002), 145--205.
\href{https://arxiv.org/abs/math/0105155}{arXiv:math/0105155}.

\bibitem{Baez2014}
J.\,C.~Baez (with contributions by G.~Egan and others),
\emph{Integral octonions} (parts 1--12),
The $n$-Category Caf\'e, November--December 2014.
The material discussed here appears in parts 9 and 10:
\url{https://golem.ph.utexas.edu/category/2014/11/integral_octonions_part_9.html},
\url{https://golem.ph.utexas.edu/category/2014/12/integral_octonions_part_10.html}.

\bibitem{ConwaySloane1999}
J.\,H.~Conway and N.\,J.\,A.~Sloane,
\emph{Sphere Packings, Lattices and Groups},
3rd edition, Springer-Verlag, 1999.

\bibitem{ConwaySmith2003}
J.\,H.~Conway and D.\,A.~Smith,
\emph{On Quaternions and Octonions: Their Geometry, Arithmetic, and
Symmetry},
A.\,K.~Peters, 2003.

\bibitem{Coxeter1946}
H.\,S.\,M.~Coxeter,
\emph{Integral Cayley numbers},
Duke Math.\ J.\ \textbf{13} (1946), 561--578.

\bibitem{Dixon2010}
G.\,M.~Dixon,
\emph{Integral octonions, octonion XY-product, and the Leech lattice},
preprint, 2010.
\href{https://arxiv.org/abs/1011.2541}{arXiv:1011.2541}.

\bibitem{Elduque1996}
A.~Elduque,
\emph{On a class of ternary composition algebras},
J.~Korean Math.\ Soc.\ \textbf{33} (1996), no.~1, 183--203.

\bibitem{Elduque2000_Triality}
A.~Elduque,
\emph{On triality and automorphisms and derivations of composition
algebras},
Linear Algebra Appl.\ \textbf{314} (2000), no.~1--3, 49--74.

\bibitem{Elduque2023_IsotropicNorm}
A.~Elduque,
\emph{Okubo algebras with isotropic norm},
in: Non-Associative Algebras and Related Topics (NAART 2020),
Springer Proc.\ Math.\ Stat., vol.~427, Springer, Cham, 2023,
pp.~287--301.
\href{https://arxiv.org/abs/2210.03456}{arXiv:2210.03456}.

\bibitem{FureyHughes2025_TrioOfTrialities}
N.~Furey and M.\,J.~Hughes,
\emph{Three generations and a trio of trialities},
Phys.\ Lett.\ B \textbf{865} (2025), 139473.
\href{https://arxiv.org/abs/2409.17948}{arXiv:2409.17948}.

\bibitem{Hall2019_Moufang}
J.\,I.~Hall,
\emph{Moufang Loops and Groups with Triality are Essentially the
Same Thing},
Memoirs of the American Mathematical Society, vol.~260, no.~1252,
Amer.\ Math.\ Soc., Providence, RI, 2019.

\bibitem{KamiyaOkubo2015}
N.~Kamiya and S.~Okubo,
\emph{Algebras satisfying triality and\/ $S_4$-symmetry},
preprint, 2015.
\href{https://arxiv.org/abs/1503.00614}{arXiv:1503.00614}.

\bibitem{Koeplinger2023}
J.~K\"oplinger,
\emph{Towards autotopies of normed composition algebras in algebraic
quantum field theory},
preprint, 2023.
\href{https://doi.org/10.13140/RG.2.2.19003.39209}{DOI:\,10.13140/RG.2.2.19003.39209}.

\bibitem{Okubo1995_Book}
S.~Okubo,
\emph{Introduction to Octonion and Other Non-Associative Algebras in
Physics},
Montroll Memorial Lecture Series in Mathematical Physics, vol.~2,
Cambridge University Press, Cambridge, 1995.

\bibitem{SmithVojtechovsky2022}
J.\,D.\,H.~Smith and P.~Vojt\v{e}chovsk\'y,
\emph{Okubo quasigroups},
Doc.\ Math.\ \textbf{27} (2022), 535--580.

\bibitem{Kirmse1924}
J.~Kirmse,
\emph{\"Uber die Darstellbarkeit nat\"urlicher ganzer Zahlen als Summen
von acht Quadraten und \"uber ein mit diesem Problem zusammenh\"angendes
nichtkommutatives und nichtassoziatives Zahlensystem},
Ber.\ Verh.\ S\"achs.\ Akad.\ Wiss.\ Leipzig, Math.-Phys.\ Kl.\
\textbf{76} (1924), 63--82.

\bibitem{Petersson2018}
H.\,P.~Petersson,
\emph{Integral octonions},
Lecture, M\'alaga Workshop on Non-Associative Algebras, 2018.

\bibitem{Wilson2009}
R.\,A.~Wilson,
\emph{Octonions and the Leech lattice},
J.~Algebra \textbf{322} (2009), 2186--2190.

\bibitem{repo}
J.~Koeplinger,
\emph{An order on the Leech lattice from transposition-twisted
octonions: research repository},
\url{https://bitbucket.org/jenskoeplinger/leech_alg}, 2026.

\end{thebibliography}

\end{document}
